# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

After unzipping the file downloaded from the web-course site.
```{r, echo=TRUE}
data <- read.csv("C:/Users/Seif/Documents/GitHub/RepData_PeerAssessment1/activity.csv")
```


## Q.1 What is mean total number of steps taken per day?
```{r}
steps_per_day <- aggregate(steps ~ date, data, sum)$steps
hist(steps_per_day)
```


### To calculate the mean:
```{r}
mean_of_data <- mean(steps_per_day)
mean_of_data
```

### To calculate the median:
```{r}
median(steps_per_day)
```

## Q.2 What is the average daily activity pattern?
```{r}
steps_by_interval <- aggregate(steps ~ interval, data, mean)
plot(steps_by_interval$steps, type="l")

```

Row of interval with highest mean steps

```{r}
max_interval = which.max(steps_by_interval$steps)
max_interval
```


## Q.3 Imputing missing values

Total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(data$steps))
```

### Devise a strategy for filling in all of the missing values in the dataset.

We can fill the missing values with the average of each day

``` {r}
## b <- steps_per_day <- aggregate(steps ~ date, data, sum)$steps
## na_indices <- which(is.na(data$steps))
## c <- steps[is.na(steps)]
## s <- split(data, data$date)

```
### I couldn't fill the values with the mean of each day, instead I will fill it with the mean of the whole finding 
```{r}
## create a duplicate, let's call it tmp for now
tmp <- read.csv("C:/Users/Seif/Documents/GitHub/RepData_PeerAssessment1/activity.csv")
tmp$steps[is.na(tmp$steps)] <- mean_of_data

```

```{r}
steps_per_day_new <- aggregate(steps ~ date, tmp, sum)$steps
hist(steps_per_day_new)
```


### To calculate the mean:
```{r}
mean_of_data_new <- mean(steps_per_day_new)
mean_of_data_new
```

### To calculate the median:
```{r}
median(steps_per_day_new)
```


### Another possible solution is to fill NA with 0
```{r}
## create another duplicate, let's call it tmp2 for now
tmp2 <- read.csv("C:/Users/Seif/Documents/GitHub/RepData_PeerAssessment1/activity.csv")
tmp2$steps[is.na(tmp2$steps)] <- 0

```

```{r}
steps_per_day_new2 <- aggregate(steps ~ date, tmp2, sum)$steps
hist(steps_per_day_new2)
```


### To calculate the mean:
```{r}
mean_of_data_new2 <- mean(steps_per_day_new2)
mean_of_data_new2
```

### To calculate the median:
```{r}
median(steps_per_day_new2)
```


### As expected the values shift to bigger values when replacing the NA with the mean of the values, and the values shift to smaller when replacing the NA with a 0 value
## Are there differences in activity patterns between weekdays and weekends?
